---
title: "Assignment 1"
author: "Mariam Mohamed"
date: "April 6, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###Include libraries
```{r}
library(dplyr)
library(ggplot2)
library(lubridate)
library(doBy)
```

###Read data
```{r}
by2ollak = read.csv("/Users/apple/Documents/Mariam/GUC/10th semester/data science/traffic-data.csv")
```

```{r}
dim(by2ollak)
str(by2ollak)
summary(by2ollak)
```


###Remove some cols
```{r}
by2ollak <- by2ollak[, !(names(by2ollak) %in% c("ad.aid", "ad.bgcl", "ad.bgcls", "ad.fncl", "ad.fncls", "ad.lid", "ad.logo", "ad.logo2x", "ad.logoAndroidS", "ad.logoAndroidH", "ad.cm", "ad.url", "ad.g", "rd.rp.type", "rd.rp.rpImg", "rd.rp.img", "rd.img", "rd.cl", "rd.rp.fullnm"))]
```

###Remove UTC from the date
```{r}
by2ollak$crawl_date = gsub(" UTC", "", by2ollak$crawl_date)
```



###Convert date to postix
```{r}
by2ollak$crawl_date = strptime(by2ollak$crawl_date, "%a %b %d %H:%M:%S %Y")
```

###Remove EET from date
```{r}
by2ollak$crawl_date = gsub(" EET", "", by2ollak$crawl_date)
```


###Add date and time cols
```{r}
by2ollak = by2ollak %>% mutate(col_month = month(crawl_date), col_day = day(crawl_date), col_wday = wday(crawl_date), col_year = year(crawl_date), col_hour = hour(crawl_date), col_minute = minute(crawl_date))
```


###Remove duplicates (same road, comment, reporter name, month, day)
```{r}
by2ollak = by2ollak[!duplicated(by2ollak[,c('rd.ri', 'rd.rp.cm', 'rd.rp.nm', 'col_month', 'col_day')]),]
```

###Data sumary
```{r}
dim(by2ollak)
str(by2ollak)
summary(by2ollak)
```

###number of reports in new 1/0 roads

```{r}
rp_roads = aggregate(by2ollak$rd.rp.cm, list(new = by2ollak$rd.new), length)
qplot(rp_roads$new, rp_roads$x, xlab = "new", ylab = "# comments")
```

###na(s) in cols
```{r}
barplot(sapply(by2ollak, function(col) sum(is.na(col))))
```



### Plots
```{r}
ggplot(by2ollak, aes(x = col_day)) + geom_histogram(binwidth = 0.5)
```

```{r}
ggplot(by2ollak, aes(x = col_hour)) + geom_histogram(binwidth = 0.5)
```

```{r}
ggplot(by2ollak, aes(x = col_wday)) + geom_histogram(binwidth = 0.5)
```

```{r}
ggplot(by2ollak, aes(x = col_month)) + geom_histogram(binwidth = 0.5)
```

###Number of comments per hour
```{r}
cm_per_hour = aggregate(by2ollak$rd.rp.cm, list(hour = by2ollak$col_hour), length)
qplot(cm_per_hour$hour, cm_per_hour$x, xlab = "hour", ylab = "# comments")

```


###Number of comments per day
```{r}
cm_per_day = aggregate(by2ollak$rd.rp.cm, list(day = by2ollak$col_day), length)
qplot(cm_per_day$day, cm_per_day$x, xlab = "day", ylab = "# comments")

```



###Number of comments per wday(wday : week day)
```{r}
cm_per_wday = aggregate(by2ollak$rd.rp.cm, list(wday = by2ollak$col_wday), length)
qplot(cm_per_wday$wday, cm_per_wday$x, xlab = "wday", ylab = "# comments")

```



###Number of comments per road
```{r}
cm_per_rd = aggregate(by2ollak$rd.rp.cm, list(road = by2ollak$rd.nm), length)
qplot(cm_per_rd$road, cm_per_rd$x, xlab = "road", ylab = "# comments")

```



###Number of comments per road and hour
```{r}
cm_per_road_hour = aggregate(by2ollak$rd.rp.cm, list(road = by2ollak$rd.nm, hour = by2ollak$col_hour), length)

```



###lazeez
```{r}
lazeez = by2ollak[grepl('^lazeez', by2ollak$rd.rp.cm),]

hist(lazeez$col_hour, xlab = "hour", breaks = c(0:24))
hist(lazeez$col_wday, xlab = "week day", breaks = c(0:7))
hist(lazeez$col_day, xlab = "day", breaks = c(0:30))
```

We can infer that people are reporting lazeez on friday ans Saturday(weekend)

###Summary of lazeez
```{r}
summary(lazeez)
```


###mashy
```{r}
mashy = by2ollak[grepl('^mashy', by2ollak$rd.rp.cm),]

hist(mashy$col_hour, xlab = "hour", breaks = c(0:24))
hist(mashy$col_wday, xlab = "week day", breaks = c(0:7))
hist(mashy$col_day, xlab = "day", breaks = c(0:30))
```

###Summary of mashy
```{r}
summary(mashy)
```


###7alawa
```{r}
halawa = by2ollak[grepl('^7alawa', by2ollak$rd.rp.cm),]

hist(halawa$col_hour, xlab = "hour", breaks = c(0:24))
hist(halawa$col_wday, xlab = "week day", breaks = c(0:7))
hist(halawa$col_day, xlab = "day", breaks = c(0:30))
```


###Summary of 7alawa
```{r}
summary(halawa)
```

###za7ma
```{r}
za7ma = by2ollak[grepl('^za7ma', by2ollak$rd.rp.cm),]

hist(za7ma$col_hour, xlab = "hour", breaks = c(0:24))
hist(za7ma$col_wday, xlab = "week day", breaks = c(0:7))
hist(za7ma$col_day, xlab = "day", breaks = c(0:30))
```

###Summary of za7ma
```{r}
summary(za7ma)
```

###mafeesh 2amal
```{r}
mafeesh_2amal = by2ollak[grepl('^mafeesh 2amal', by2ollak$rd.rp.cm),]

hist(mafeesh_2amal$col_hour, xlab = "hour", breaks = c(0:24))
hist(mafeesh_2amal$col_wday, xlab = "week day", breaks = c(0:7))
hist(mafeesh_2amal$col_day, xlab = "day", breaks = c(0:30))
```

###Summary of mafeesh_2amal
```{r}
summary(mafeesh_2amal)
```


###Models
```{r}
model1 <- lm(col_hour ~ col_wday, data=by2ollak)
summary(model1)
layout(matrix(c(1,2,3,4),2,2)) # optional 4 graphs/page 
plot(model1)
termplot(model1)
```

```{r}
coplot(col_wday~col_day|col_hour, data = by2ollak)
```

```{r}
model2 = lm(col_hour~col_day + col_wday, data = by2ollak)
layout(matrix(c(1,2,3,4),2,2)) # optional 4 graphs/page 
plot(model2)
summary(model2)
```

```{r}
anova(model1, model2)
```


```{r}
model3 <- lm(col_hour ~ col_wday, data=lazeez)
summary(model3)
layout(matrix(c(1,2,3,4),2,2)) 
plot(model3)
termplot(model3)
```

```{r}
summaryBy(rd.nm ~ col_hour + col_wday, data = by2ollak, 
 	FUN = function(x) { c(num_of_roads = length(unique(x))) } )
```

